---
title: "Data608-HW1"
author: "Armenoush Aslanian-Persico"
output:
  pdf_document: default
  html_document: default
---

Principles of Data Visualization and Introduction to ggplot2

## 1. Show Companies by States.

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.1. Load Libraries

```{r libraries, warning=FALSE, message=FALSE}
library(DT)
library(ggplot2)
library(jsonlite)
library(knitr)
library(plyr)
library(dplyr)
library(RCurl)
library(reshape2)
library(rmarkdown)
library(stringr)
library(tidyr)
```


## 1.2. Load Data from URL

```{r loaddata}
companiescsv<-getURL("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA608/master/lecture1/Data/inc5000_data.csv")

df1<- data.frame(read.csv(text=companiescsv))

datatable(df1, options = list( pageLength = 5, lengthMenu = c(5, 10),  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#00838f', 'color': '#fff'});", "}"), rownames=TRUE))

```


## 1.3. Group Companies by State

```{r groupbystate}

#show data structure
str(df1)

#group companies by state
groupbystate <- ddply(df1, .(State), summarize, count = length(Rank))

orderbystate <- groupbystate[order(groupbystate$count), ]

#order by state with most companies
orderbystate$State <- factor(orderbystate$State, levels = orderbystate$State)  
```


## 1.4 Show Graph of Companies by State.

```{r graphbystate}
#Create graph of companies by state
Figure1 <- 
  ggplot(orderbystate, aes(x = State, y = count)) + 
  geom_bar(stat = "identity", fill="#ffca28")+
  coord_flip() + 
  theme_bw()+
  ggtitle("Number of Fastest Growing US Companies by State") + 
  labs(x="State",y="Number of Companies")
  
Figure1
ggsave("Figure1.jpg")
```

## 2. Show Average Employment by Industry in the Third Ranked State.

Let’s dig in on the State with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries employ. Create a plot of average employment by industry for companies in this state (only use cases with full data (user R’s complete.cases() function). Your graph should show how variable the ranges are, and exclude outliers.

## 2.1 Sort States by Count of Companies

```{r sortstates}
allstates <- as.data.frame(table(df1$State))
colnames(allstates) <- c("State", "Count")

# sort by descending count
sortstates <- allstates[order(-allstates$Count),] 
head(sortstates)

```

## 2.2 Select State with Third Most Companies

```{r selectstate}
#subset third highest value
x <- sort(allstates$Count, TRUE)[3] 
filter(allstates, Count == x)

#filter by resulting state
nys <- filter(df1, State == "NY")

#subset complete cases only
nys <- nys[complete.cases(nys),]

#show data structure
str(nys)

```

## 2.3 Calculate Average Employment by Industry

```{r nysjobs, eval=TRUE}

#Create summary columns for mean and count
nysjobs <- ddply(nys, .(Industry), summarize,
    meanemployees = round(mean(Employees),0), 
    sumemployees = sum(Employees), 
    countemployers = length(Employees), 
    meanrevenueperemployee = round(mean(Revenue/Employees),2),
    meangrowthrate = round(mean(Growth_Rate),2)
    )

#Show industries with highest average employee count
nysjobs <- nysjobs[order(nysjobs$meanemployees, decreasing = TRUE),]

#Show calculated data
datatable(nysjobs, options = list( pageLength = 5, lengthMenu = c(5, 10),  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#1565c0', 'color': '#fff'});", "}"), rownames=TRUE))

#Create industry factor
nys$Industry <- factor(nys$Industry, levels = nysjobs$Industry)

#Show nys data
datatable(nys, options = list( pageLength = 5, lengthMenu = c(5, 10),  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#283593', 'color': '#fff'});", "}"), rownames=TRUE))
```


## 2.4 Show Graph of Average Employment by Industry.

```{r graphnysjobs}

#Create graph of jobs by industry
Figure2 <-
ggplot(nys, aes(x = Industry, y = Employees)) +
  geom_boxplot() + 
  stat_summary(fun.y = mean, geom = "point", shape = 20, size = 5, color = "blue") + 
  theme_bw()+
  ggtitle("New York State Average Employment by Industry")

#select limits to hide outliers
Figure2<- Figure2 + coord_flip(ylim = c(0,1500))
Figure2
ggsave("Figure2.jpg")
```

## 3. Show Highest Revenue Industries.

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart makes this information clear.

```{r revenuejobs}
nysjobs1 <- nysjobs[order(nysjobs$meanrevenueperemployee, decreasing = TRUE),]

#Create graph of jobs by revenue
Figure3 <-
ggplot(nysjobs1, aes(x = reorder(Industry,-meanrevenueperemployee), y = meanrevenueperemployee)) +
  geom_bar(stat = "identity", fill="#ffca28")+
  theme_bw()+
  ggtitle("New York State Average Employee Revenue by Industry")+
  labs(y="Revenue in Dollars",x="Industry")

Figure3<- Figure3 + coord_flip()
Figure3
ggsave("Figure3.jpg")
```