---
title: "opendata"
author: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r 0-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(data.table)
library(DT)
library(openxlsx)
library(dtplyr)
library(dplyr)
library(base)
library(ggplot2)
library(reshape2)
```

```{r 1-scrap, eval=FALSE, echo=FALSE}
#########################################################################
#START SCRAP
########################################################################

#df<-read.csv("data1.csv", header=TRUE, sep=",")
#data.table(head(df))
#df<- data.frame(read.table("cro2.xlsx", header=TRUE, fill=TRUE))
#df <-read.csv("cro2.csv", header=TRUE, stringsAsFactors=FALSE)
#df<-(head(df,10))

#group raw data by agency  #test
graw<-(df %>%  group_by(AgencyName) %>% summarise(n = n()))
#display in table
datatable(graw, options = list( pageLength = 10, lengthMenu = c(10, 20, 40),   initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#01975b', 'color': '#fff'});",
    "}")), rownames=TRUE)

#group by agency 
g2<-(df1 %>% 
  group_by(AgencyName) %>% 
  summarise(n = n(), mean=mean(ContractAmount))

datatable(g2, options = list( pageLength = 10, lengthMenu = c(10, 20, 40),   initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#01975b', 'color': '#fff'});",
    "}")), rownames=TRUE)
#########################################################################
#END SCRAP
########################################################################
```

```{r 1-load-clean-data, eval=FALSE, echo=FALSE}
########################################################################
#START EVAL=FALSE
########################################################################

#import data
df <- read.xlsx("cro2.xlsx", sheet=1, startRow=1, colNames=TRUE)

#format dates
df$StartDate<- as.Date(df$StartDate, origin = "1899-12-30")
df$StartDate<- format(df$StartDate, "%Y")
df$EndDate<- as.Date(df$EndDate, origin = "1899-12-30")
df$EndDate<- format(df$EndDate, "%Y")

#replace miscoded characters; combine agencies with multiple names
df$AgencyName[df$AgencyName == "ADMIN FOR CHILDREN'S SVCS" | df$AgencyName == "ADMIN FOR CHILDREN_x001A_S SVCS" | df$AgencyName == "Administration for Children's Services"] <- "ACS"
df$AgencyName[df$AgencyName =="HRA/DEPT OF SOCIAL SERVICES"]<-"Human Resources Administration"
df$AgencyName[df$AgencyName =="Citywide Administrative Services"]<-"DEPT OF CITYWIDE ADMIN SVCS"
df$AgencyName[df$AgencyName =="Parks and Recreation"]<-"DEPT OF PARKS & RECREATION"

#Selected top 13 agencies by activity volume, excluding elections
df1<-df[df$AgencyName %in% c("POLICE DEPARTMENT", 
                        "DEPARTMENT OF EDUCATION ADMIN",
                        "DEPARTMENT OF SANITATION",
                        "DEPT OF PARKS & RECREATION",
                        "Human Resources Administration",
                        "FIRE DEPARTMENT",
                        "DEPARTMENT OF CORRECTION",
                        "DEPT OF HEALTH/MENTAL HYGIENE",
                        "ACS",
                        "DEPT OF ENVIRONMENT PROTECTION",
                        "DEPARTMENT OF TRANSPORTATION",
                        "DEPT. OF HOMELESS SERVICES",
                        "DEPT OF CITYWIDE ADMIN SVCS"
), ]

#Selected Changes in Personnel, Contract Award, Procurement notices only
df1<-df1[df1$SectionName %in% c("Changes in Personnel", "Contract Award Hearings", "Procurement"),]

#export cleaned data
write.csv(df1, file = "df1.csv)"
########################################################################
#END EVAL=FALSE
########################################################################
```

```{r 2-group-agency, echo=FALSE}

#import cleaned data
df1 <- read.xlsx("df.xlsx", sheet=1, startRow=1, colNames=TRUE)

#group by agency and action type
t1<-(df1 %>% 
  group_by(Agency, SectionName,Year) %>% 
  summarise(n = n() ))
```

## Data Cleanup: Selected Agencies Only, Grouped by Year

The raw data was grouped by year.
<br>

Selected agencies:
<br>
/// List of selected agencies and descriptions

```{r 3-check-str-1, eval=FALSE, echo=FALSE}
#check structure
str(df1)
```

## Show Table: Selected Agencies, All Actions, Grouped by Year

```{r 4-show-agency, eval=TRUE, echo=FALSE}
#create table to display data
datatable(t1, options = list( pageLength = 10, lengthMenu = c(10, 20, 40),   initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#01975b', 'color': '#fff'});",
    "}")), rownames=TRUE)

t1<-as.data.frame(t1)
```

## Show Table: Selected Agencies, All Actions, Columns by Year
### The dataset was reshaped to show years as columns.

```{r 5-graphbyyear, eval=TRUE, echo=FALSE}
m2 <- dcast(t1, Agency + SectionName ~ Year)

m1<-as.data.table(m2)

datatable(m1, options = list( pageLength = 10, lengthMenu = c(10, 20, 40),   initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#01975b', 'color': '#fff'});",
    "}")), rownames=TRUE)
```

## Show Table: Selected Agencies, Hiring Actions, Columns by Quarter

```{r 6-import-summary-data, echo=FALSE}

#import cleaned data
df2 <- read.xlsx("df-agencybyqtr.xlsx", sheet=1, startRow=1, colNames=TRUE)

#group by agency and action type

datatable(df2, options = list( pageLength = 10, lengthMenu = c(10, 20, 40),   initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#3498DB', 'color': '#fff'});",
    "}")), rownames=TRUE)
```


